---
title: "ITMS 514 Final Project"
author: "Sarvesh Kaushik, Jay Gupta, Colin Brennan"
date: "4/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include=FALSE}

library(ggplot2)
library(ISLR)
library(tidyverse)
library(dplyr)
library(tigerstats)
library(MLmetrics)
library(MASS)

```

**Input of both region's CSV**
```{r}
forestfireregion1 <- read.csv("region1.csv", sep = ",") 
forestfireregion2 <- read.csv("region2.csv", sep = ",") 
```


**Data Preparation Region 1**
```{r}
dim(forestfireregion1)
str(forestfireregion1)
summary(forestfireregion1)
head(forestfireregion1)
sum(is.na(forestfireregion1))
sum(duplicated(forestfireregion1))
```

**Data Preparation Region 2**
```{r}
dim(forestfireregion2)
str(forestfireregion2)
summary(forestfireregion2)
head(forestfireregion2)
sum(is.na(forestfireregion2))
sum(duplicated(forestfireregion2))
```

**Bar Plot Region 1**
```{r}
# Fires Verse Not Fire 
ggplot(data = forestfireregion1) + geom_bar(mapping = aes(x = Classes))

#Fires compared with Wind Speed
ggplot(data = forestfireregion1, aes(x = Ws)) +
  geom_bar(aes(fill = Classes)) 
```

**EDA Region 1**
```{r}
ws1=forestfireregion1$Ws
h1=hist(ws1,ylim=c(0,40))
text(h1$mids,h1$counts,labels=h1$counts, adj=c(0.5, -0.5))
hist(ws1,main="Maximum daily Wind Speed",xlab="Wind Speed",xlim=c(0,30),col="darkmagenta",freq=FALSE)
boxplot(ï..day~Ws,data=forestfireregion1,main="Measure of wind speed per day",xlab="Wind Speed",ylab="Days")
boxplot(Rain~Ws,data=forestfireregion1,main="Measure of rain according to the wind speed",xlab="Wind Speed",ylab="Rain")
ggplot(data = forestfireregion1, aes(x = Ws)) +geom_histogram(aes(fill = Classes),bins = 30)
```


**EDA Region 2**
```{r}
ws2=forestfireregion2$Ws
h2=hist(ws2,ylim=c(0,70))
text(h2$mids,h2$counts,labels=h2$counts, adj=c(0.5, -0.5))
hist(ws2,main="Maximum daily Wind Speed",xlab="Wind Speed",xlim=c(0,30),col="darkmagenta",freq=FALSE)
boxplot(ï..day~Ws,data=forestfireregion2,main="Measure of wind speed per day",xlab="Wind Speed",ylab="Days")
boxplot(Rain~Ws,data=forestfireregion2,main="Measure of rain according to the wind speed",xlab="Wind Speed",ylab="Rain")
ggplot(data = forestfireregion2, aes(x = Ws)) +geom_histogram(aes(fill = Classes),bins = 30)
```


**Count Plot Fire Vs WS Region 1**
```{r}
ggplot(data = forestfireregion1) +
  geom_count(mapping = aes(x = Classes, y = Ws))
```

**Count Plot Fire Vs WS Region 2**
```{r}
ggplot(data = forestfireregion2) +
  geom_count(mapping = aes(x = Classes, y = Ws))
```


**Histogram RH Classes Region 1**
```{r}
ggplot(data = forestfireregion1, aes(x = RH)) +
  geom_histogram(aes(fill = Classes), bins =10, color = "black")
```

**Histogram RH Classes Region 2**
```{r}
ggplot(data = forestfireregion2, aes(x = RH)) +
  geom_histogram(aes(fill = Classes), bins =10, color = "black")
```



**Geom Point Region 1**
```{r}
ggplot(data = forestfireregion1) +
  geom_point(mapping = aes(x = Rain, y = Temperature,colour = Classes))
```

**Geom Point Region 2**
```{r}
ggplot(data = forestfireregion2) +
  geom_point(mapping = aes(x = Rain, y = Temperature,colour = Classes))
```


## Multiple Regression testing Region 1
```{r}
set.seed(420)

i <- sample(2, nrow(forestfireregion1), replace=TRUE, prob=c(0.8, 0.2))

region1Training <- forestfireregion1[i==1,]

region1Testing <- forestfireregion1[i==2,]

fitlm <- lm(Ws~., data=region1Training[,1:14])
summary(fitlm)

fitlm2<- lm(Ws~.-year-month-RH, data=region1Training)
summary(fitlm2)

ypred <-predict(object = fitlm2, newdata = region1Testing[,1:14])
summary(ypred)
```

## Multiple Regression testing Region 2
```{r}
set.seed(420)
i2 <- sample(2, nrow(forestfireregion2), replace=TRUE, prob=c(0.8, 0.2))

region2Training <- forestfireregion2[i2==1,]

region2Testing <- forestfireregion2[i2==2,]


fitlm3 <- lm(Ws~., data=region2Training[,1:14])
summary(fitlm3)

fitlm4<- lm(Ws~.-year-month-RH, data=region2Training)
summary(fitlm4)

ypred2 <-predict(object = fitlm4, newdata = region2Testing[,1:14])
summary(ypred2)

```

**Forward Step Regression Region 1**
```{r}
intercept_only <- lm(Ws ~ 1, data=region1Training[,1:14])
all <- lm(Ws~., data=region1Training[,1:14])
forward <- stepAIC (intercept_only, direction='forward',scope = formula(all))
forward$anova
ypred_forward <-predict(object = forward, newdata = region1Training[,1:14])
MAE(y_pred = ypred_forward, y_true = region1Testing$Ws)
MSE(y_pred = ypred_forward, y_true =  region1Testing$Ws)
```

**Forward Step Regression Region 2**
```{r}
intercept_only <- lm(Ws ~ 1, data=region2Training[,1:14])
all <- lm(Ws~., data=region2Training[,1:14])
forward2 <- stepAIC (intercept_only, direction='forward',scope = formula(all))
forward2$anova
ypred_forward2 <-predict(object = forward2, newdata = region2Training[,1:14])
MAE(y_pred = ypred_forward2, y_true = region2Testing$Ws)
MSE(y_pred = ypred_forward2, y_true =  region2Testing$Ws)

```

**Backwards Selection**

```{r}
backward <- stepAIC (all, direction='backward')
```



```{r}
backward$anova
```

```{r}
summary(backward)
```
