---
title: "Project"
author: "Jay Gupta"
date: '2022-04-16'
output: html_document
---

---
title: "Algerian_forest_fires_Project"
author: "Jay Gupta"
date: '2022-04-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ISLR)
```

#Reading the csv file
```{r}
forestfireregion1=read.csv("/Users/jaygupta/Downloads/Forest FireRegion1.csv",sep=",")
forestfireregion2=read.csv("/Users/jaygupta/Downloads/ForestFireRegion2.csv",sep=",")
```

#Viewing the csv file
```{r}
View(forestfireregion1)
View(forestfireregion2)
```

#Data Preparation
```{r}
dim(forestfireregion1)
dim(forestfireregion2)
str(forestfireregion1)
str(forestfireregion2)
summary(forestfireregion1)
summary(forestfireregion2)
head(forestfireregion1)
head(forestfireregion2)
sum(is.na(forestfireregion1))
sum(is.na(forestfireregion2))
sum(duplicated(forestfireregion1))
sum(duplicated(forestfireregion2))
```

#EDA
```{r}
temperature1=forestfireregion1$Temperature
temperature2=forestfireregion2$Temperature
h1=hist(temperature1,ylim=c(0,40))
text(h1$mids,h1$counts,labels=h1$counts, adj=c(0.5, -0.5))
h2=hist(temperature2,ylim=c(0,40))
text(h2$mids,h2$counts,labels=h2$counts, adj=c(0.5, -0.5))
hist(temperature1,main="Maximum daily temperature",xlab="Temperature in degrees Fahrenheit",xlim=c(0,40),col="darkmagenta",freq=FALSE)
boxplot(day~Temperature,data=forestfireregion1,main="Measure of temperature per day",xlab="Temperature",ylab="Days")
boxplot(rain~Temperature,data=forestfireregion1,main="Measure of rain according to the temperature",xlab="Temperature",ylab="Rain")
boxplot(day~Temperature,data=forestfireregion2,main="Measure of temperature per day",xlab="Temperature",ylab="Days")
boxplot(rain~Temperature,data=forestfireregion2,main="Measure of rain according to the temperature",xlab="Temperature",ylab="Rain")
ggplot(data = forestfireregion1, aes(x = Temperature)) +geom_histogram(aes(fill = Classes),bins = 30)
ggplot(data = forestfireregion2, aes(x = Temperature)) +geom_histogram(aes(fill = Classes),bins = 30)
```
>breaks-places where the breaks occur,
counts-the number of observations falling in that cell,
density-the density of cells, mids-the midpoints of cells,
xname-the x argument name and
equidist-a logical value indicating if the breaks are equally spaced or not.

#Data Analysis

#Hypothesis Testing:
```{r}
junemonth=subset.data.frame(forestfireregion1,forestfireregion1$month==6)
summary(junemonth)
julymonth=subset.data.frame(forestfireregion1,forestfireregion1$month==7)
summary(julymonth)
s1=sd(junemonth$Rain)
s2=sd(julymonth$Rain)
n1=length(junemonth$Rain)
n2=length(julymonth$Rain)
fvalue=s1^2/s2^2
fvalue
alpha=0.05
f1=qf(alpha/2,n1-1,n2-1)
f1
f2=qf(1-alpha/2,n1-1,n2-1)
f2
var.test(julymonth$Rain,junemonth$Rain,alternative = "two.sided",conf.level = 0.95)
```
##Hence the variances are not equal because the p value is much lesser than significance level and the fvalue doesn't lie between f1 and f2. H0: var(junemonth$rain)=var(julymonth$rain). H1: var(junemonth$rain)≠var(julymonth$rain)

##For unknown mean and unequal variances

```{r}
t.test(junemonth$Rain,julymonth$Rain,var.equal = FALSE,conf.level = 0.95)
```
##We are using t.test because we have two means and the variance are unknown and are not equal. The p-value is greater than alpha i.e., 0.05. So we can accept the null hypothesis H_0 and agree that the increase of total rain in the month of June is equal to the increase of total rain in the month of July.

#Regression:

```{r}
fitlm=lm(Rain~Classes,data=forestfireregion1)
summary(fitlm)
```

```{r}
i <- sample(2, nrow(forestfireregion1), replace=TRUE, prob=c(0.8, 0.2))
ForestfireTraining <- forestfireregion1[i==1,]
ForestfireTesting <- forestfireregion1[i==2,]
pairs(ForestfireTraining[,1:13],lower.panel = NULL)
```
```{r}
cor(subset(ForestfireTraining,select=-Classes))
```

```{r}
fitlm=lm(Temperature~.-year,data=ForestfireTraining[,1:13])
summary(fitlm)
```
#By performing 